---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# RegularizedFits

The goal of RegularizedFits is fit polynomials to data in a way that avoids overfitting.

## Example

Here is a basic example that demonstrates its applicability. If we have the data:

```{r}
x <- 0:5
y <- (0:5)^2 + c(-0.40, 3.50, 1.80, 10.00, -9.75, 13.80)
data <- data.frame(y,x) 
```


Then this is the result of cubic fits using `lm()`, `reg.lm()`, and `crossval()`

```{r}
#lm
fit <- lm(y ~ cbind(x, x^2,x^3))
fit <- as.numeric(fit$coefficients)
fit
```

We can see that all the coefficients are quite large leading to a lot of wavering of the function.

```{r}
#reg.lm
regcube <- reg.lm(y, x, degree = 3)
regcube
regfit <- regcube$coefficients
```

Here we can see a bit more moderation in the coefficients, especially in the 3rd order one. Note that this one used lambda = 1.

We can see if there's any values of lambda that give better fits by using `cv.fit()`:

```{r}
#crossval
cv.fit <- crossval(y, x, degree = 3, lambda = c(0.1, 1, 10))
cv.fit
cv.fit <- cv.fit$coefficients
```

This one has even more moderation. Note that although the error here is larger than for `reg.lm()`; this doesn't mean that `crossval()` gave a worse fit. This happens because `crossval()` looks at the ability to make predictions of general trends, which if you plot you will see that its fit is closest to the x^2 function to the naked eye in the following plot:

```{r}
ggplot(data, aes(x = x, y = y)) + geom_point() + 
	stat_function(fun = function(x) fit[1] + fit[2]*x + fit[3]*x^2 + fit[4]*x^3, aes(color = 'unregularized')) + 
	stat_function(fun = function(x) regfit[1] + regfit[2]*x + regfit[3]*x^2 + regfit[4]*x^3, aes(color = 'regularized')) + 	
	stat_function(fun = function(x) cv.fit[1] + cv.fit[2]*x + cv.fit[3]*x^2 + cv.fit[4]*x^3, aes(color = 'cross validated')) +
	stat_function(fun = function(x) x^2, color = 'grey') +
	theme(legend.position = "bottom") + theme_bw()
```

For more detail, see the vignette.

